### 2021-04-12 paper/natastron/ArXiV/paper.tex
prepped the ArXiV submission

\usepackage{chngcntr}
\usepackage{xspace}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{lscape}
\usepackage{longtable}
\usepackage{threeparttable}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[ ]{hyperref}
%\graphicspath{{Images/}}
%\usepackage[symbol]{footmisc}

@@

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
 %500 words
Gyrochronology is the study of the relationship between a star's rotation period and its age. As a star grows older along the main sequence, magnetic winds will cause it to lose angular momentum, slowing its rotation. The loss rate is thought to depend on the depth of the convection zone, which is a strong function of mass and temperature, and so the rotation period of a young star will rapidly settle on to a plane in age-colour-rotation space \cite{barnes2007}. Knowing the rotation and colour of stars therefore provides an avenue to measure ages, to a precision of $\sim 10\%$ for stars most similar to the Sun \cite{meibom+2015}. Age can otherwise be difficult to infer, and so gyrochronology enables more in-depth studies of stellar populations and individual stars \cite{leiner+2019,claytor+2019}.

@@

In order to ensure our results were based on the most robust rotation measurements, we focused on the stars with the best convergence metrics in the mode frequency fitting, excluding those with a number of effective samples $n_{\rm eff} < 1000$ and a Gelman-Rubin metric of $\hat{R}<1.1$ \cite{gelman+rubin1992,salvatier+2016} (see  \textbf{Supplementary Information} Table 1). Two stars with low metallicities ($< -0.8\, \rm {dex}$) that fell outside the boundaries of our stellar models were also excluded. The remaining sample of 73 stars contained 4 potential sub-giants, 22 `hot' stars and 47 main sequence stars. Figure \ref{fig:fullsample} shows these stars plotted over the top of both the standard and WMB population models in rotation-temperature space, alongside the stars not used in the gyrochronology analysis (shown without uncertainties).

@@

From this asteroseismic analysis, we report inclination angle ($i$), rotational splitting ($\nu_{\rm s}$), and rotation period ($P$) in \textbf{Supplementary Information} Table 1. The summary statistics on these parameters were taken as the median of the posterior distribution, with uncertainties being the $15.9^{\rm th}$ and $84.1^{\rm st}$ percentiles. For inclination angle ($i = \arccos(\cos(i))$) and rotation ($P = 1/\nu_{\rm s}$), the full posterior samples were transformed before taking the summary statistics, as our model sampled in $\cos(i)$ and $\nu_{\rm s}$.

We flagged any sub-optimal conditions of the final fit. We flagged 5 for which the Gelman-Rubin convergence metric, $\hat{R}$, was greater than 1.01 and 2 stars for which it was greater than 1.1 on inclination angle or rotation, where a $\hat{R} = 1$ indicates a converged result \cite{gelman+rubin1992}. We also performed visual checks of the sampled chains on all hyperparameters and of the best-fit model compared to both the raw and smoothed asteroseismic data. We found no issues in the visual investigation of 94 stars. KIC 8478994 is not reported due to both a poor unconverged fit as well as high $\hat{R}$ on rotational parameters. KICs 6603624, 8760414 and 8938364 are reported in  \textbf{Supplementary Information} Table 1, but were excluded from the gyrochronology analysis due to strong disagreement with multiple studies in the literature (see main body of paper). Finally, we flagged any stars with fewer than 1000 effective samples ($n_{\rm eff}$) of $\nu_s$.

@@

\textbf{Data availability statement:} The core input data and results are summarised in  \textbf{Supplementary Information} Table 1, which is published with this article, and is also available on Vizier. Larger data files, such as stellar model populations and individual posterior distribution chains from the asteroseismic and gyrochronology model fitting are fully available on request.

@@

\setcounter{figure}{0}    
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6 \textwidth]{Images/pgm_models.pdf}
	\caption{\textbf{A probabilistic graphical model (PGM) represented algebraically in Equation \ref{eq:modelll}.} The shaded circle indicates observed data, and solid black points represent other fixed information, such as the KDEs and observational uncertainties. The remaining circles represent parameters. The underline indicates that the symbol represents a set of parameters or data. Here, $\kappa_{\rm{s}}$ and $\kappa_{\rm WMB}$ represent the KDEs of standard and WMB model populations respectively. $Q_{\rm{WMB}}$ is the mixture model weighting factor. The latent parameters $\theta$, our observations $\mathcal{D}$ and their uncertainties $\sigma_{\mathcal{D}}$ include temperature (\teff), mass ($M$), log-age ($\ln(t)$), metallicity (\feh) and log-rotation ($\ln(P)$). This model is \textit{hierarchical}, as all the latent parameters are drawn from the common probability distribution set by $Q_{\rm{WMB}}$ and described in Equation \ref{eq:mixturell}.}
	\label{fig:pgm}
\end{figure}

>>>>

\usepackage{chngcntr}
\usepackage{xspace}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{lscape}
\usepackage{longtable}
\usepackage{threeparttable}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[ ]{hyperref}
\graphicspath{{Images/}}
%\usepackage[symbol]{footmisc}

@@

%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%
 %500 words
Gyrochronology is the study of the relationship between a star's rotation period and its age. As a star grows older along the main sequence, magnetic winds will cause it to lose angular momentum, slowing its rotation. The loss rate is thought to depend on the depth of the convection zone, which is a strong function of mass and temperature, and so the rotation period of a young star will rapidly settle on to a plane in age-colour-rotation space \cite{barnes2007}. Knowing the rotation and colour of stars therefore provides an avenue to measure ages, to a precision of $\sim 10\%$ for stars most similar to the Sun \cite{meibom+2015}. Age can otherwise be difficult to infer, and so gyrochronology enables more in-depth studies of stellar populations and individual stars \cite{leiner+2019,claytor+2019}.

@@

In order to ensure our results were based on the most robust rotation measurements, we focused on the stars with the best convergence metrics in the mode frequency fitting, excluding those with a number of effective samples $n_{\rm eff} < 1000$ and a Gelman-Rubin metric of $\hat{R}<1.1$ \cite{gelman+rubin1992,salvatier+2016} (see  \textbf{Supplementary Table 1}). Two stars with low metallicities ($< -0.8\, \rm {dex}$) that fell outside the boundaries of our stellar models were also excluded. The remaining sample of 73 stars contained 4 potential sub-giants, 22 `hot' stars and 47 main sequence stars. Figure \ref{fig:fullsample} shows these stars plotted over the top of both the standard and WMB population models in rotation-temperature space, alongside the stars not used in the gyrochronology analysis (shown without uncertainties).

@@

From this asteroseismic analysis, we report inclination angle ($i$), rotational splitting ($\nu_{\rm s}$), and rotation period ($P$) in \textbf{Supplementary Table 1}. The summary statistics on these parameters were taken as the median of the posterior distribution, with uncertainties being the $15.9^{\rm th}$ and $84.1^{\rm st}$ percentiles. For inclination angle ($i = \arccos(\cos(i))$) and rotation ($P = 1/\nu_{\rm s}$), the full posterior samples were transformed before taking the summary statistics, as our model sampled in $\cos(i)$ and $\nu_{\rm s}$.

We flagged any sub-optimal conditions of the final fit. We flagged 5 for which the Gelman-Rubin convergence metric, $\hat{R}$, was greater than 1.01 and 2 stars for which it was greater than 1.1 on inclination angle or rotation, where a $\hat{R} = 1$ indicates a converged result \cite{gelman+rubin1992}. We also performed visual checks of the sampled chains on all hyperparameters and of the best-fit model compared to both the raw and smoothed asteroseismic data. We found no issues in the visual investigation of 94 stars. KIC 8478994 is not reported due to both a poor unconverged fit as well as high $\hat{R}$ on rotational parameters. KICs 6603624, 8760414 and 8938364 are reported in  \textbf{Supplementary Table 1}, but were excluded from the gyrochronology analysis due to strong disagreement with multiple studies in the literature (see main body of paper). Finally, we flagged any stars with fewer than 1000 effective samples ($n_{\rm eff}$) of $\nu_s$.

@@

\textbf{Data availability statement:} The core input data and results are summarised in  \textbf{Supplementary Table 1}, which is published with this article, and is also available on Vizier. Larger data files, such as stellar model populations and individual posterior distribution chains from the asteroseismic and gyrochronology model fitting are fully available on request.

@@

\renewcommand{\figurename}{Extended Data Figure}

\setcounter{figure}{0}    
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6 \textwidth]{pgm_models.pdf}
	\caption{\textbf{A probabilistic graphical model (PGM) represented algebraically in Equation \ref{eq:modelll}.} The shaded circle indicates observed data, and solid black points represent other fixed information, such as the KDEs and observational uncertainties. The remaining circles represent parameters. The underline indicates that the symbol represents a set of parameters or data. Here, $\kappa_{\rm{s}}$ and $\kappa_{\rm WMB}$ represent the KDEs of standard and WMB model populations respectively. $Q_{\rm{WMB}}$ is the mixture model weighting factor. The latent parameters $\theta$, our observations $\mathcal{D}$ and their uncertainties $\sigma_{\mathcal{D}}$ include temperature (\teff), mass ($M$), log-age ($\ln(t)$), metallicity (\feh) and log-rotation ($\ln(P)$). This model is \textit{hierarchical}, as all the latent parameters are drawn from the common probability distribution set by $Q_{\rm{WMB}}$ and described in Equation \ref{eq:mixturell}.}
	\label{fig:pgm}
\end{figure}
